---
title: "modComp study analysis of causal attribution task data"
output:
  html_document:
    html-math-method:
      method: mathjax
  #  pdf_document:
  # extra_dependencies: ["bbm"]
  # fig_caption: yes
---

```{r setup, include=FALSE}
# knitr options
knitr::opts_chunk$set(echo=FALSE, error=TRUE, warning=FALSE, message=FALSE, fig.align='center')

# load packages
packages <- c("rstan", "dplyr", "tidyr", "bayesplot", "loo", "hBayesDM", "tidybayes", "forcats",
              "ggpmisc", "patchwork", "devtools", "reshape2", "ggExtra", "unikn", "svglite",
              "lme4", "glmnet", "ggcorrplot", "subscore", "rstatix", "boot")
if (length(setdiff(packages, rownames(installed.packages()))) > 0 ) {
  install.packages(setdiff(packages, rownames(installed.packages())))
}
lapply(packages, require, character.only=TRUE)

# set wd
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

# what task version to filter for
task_ver <- "causal-attr-1-2"

# # create figures subdir
# subdir <- "figures"
# if (!file.exists(subdir)){
#   dir.create(file.path(dirname(rstudioapi::getActiveDocumentContext()$path), subdir))
# }
```

```{r setup_rstan}
rstan_options(auto_write = TRUE)     # write the models so we don't have to recompile each time
nCores <- parallel::detectCores()    # get number of cores available for parallelisation
```

```{r setup_colour_scales}
# lets set some custom colour scales for our plots using unikn
#seecol(pal_unikn_pair)
palette2 <- usecol(pal_unikn_pair) 
colours3 <- c("restructuring" = palette2[2],
              "control" = palette2[3]               # restructuring intervention control
              )
# colours by model parameter type
colours2 <- c("group mean" = palette2[14],
              "intervention effect" = palette2[2]   # here = restructuring
              )
```

```{r load_data}
# load data
data_long_all <- read.csv(file=paste0(task_ver, "-causal-attribution-task-data-anon.csv")) %>%
  dplyr::select(-nTrials, -ID, -X) %>%
  arrange(uid, taskNo, itemNo) %>%
  mutate(sess = taskNo + 1,
         neg_pos = ifelse(valence=="negative",0,1))

## get number of time points etc
nPpts <- length(unique(data_long_all$uid))
nTimes <- max(data_long_all$sess)
nTrials_all <- data_long_all %>%
  arrange(uid) %>%
  group_by(uid) %>%
  summarize(nTrials = n()) %>%
  mutate(ID = seq(1, nPpts, 1))          # assign sequential numeric IDs for later
data_long_all <- merge(data_long_all, nTrials_all, by="uid")

nTrials_max <- nTrials_all %>%
  {max(.$nTrials)}

# get lists of subjects IDs by condition for use with other data
control_subs <- data_long_all %>%
  filter(condition=="control") %>%
  dplyr::select(uid, ID)
controls <- as.list(unique(control_subs$uid))
control_IDs <- as.list(unique(control_subs$ID))

# get ordered list of intervention conditions
int_conds <- data_long_all %>%
  arrange(ID) %>%
  group_by(ID) %>%
  dplyr::select(ID, condition) %>%
  distinct() %>%
  mutate(condition01 = ifelse(condition=="psychoed", 1, 0))
```

### Via generative model (both sessions data)

Latent traits of tendency to endorse an internal and global cause are modelled separately (bernoulli function with single parameter theta governing probability of endorsement on each trial)

Since estimates of theta were previously observed to be more precise when separate values were allowed for positive and negative events, we again allow different parameters for different valences.

Values of theta for each session were modelled separately but were assumed to be drawn from a multivariate normal distribution, i.e., were allowed to covary across sessions, and between internal and global attribution tendencies within sessions (uniform prior ranging -1 to 1), allowing direct estimation of correlation of estimates between sessions (as per [Rouder et al., 2019](https://doi.org/10.3758/s13423-018-1558-y) and [Haines et al., (2020)](https://psyarxiv.com/xr7y3/)).

```{r stan_int_models_IG}
## specify model, session, data to fit, and params to save
model <- "m_bernoulli_negpos_IGcorr2_multisess_intervention_additive"

## create arrays of choice options and responses for each participant and time point
internalChosen_neg <- internalChosen_pos <- globalChosen_neg <- globalChosen_pos <- array(0, dim = c(nPpts, nTimes, nTrials_max/2))
nT_ppts <- array(nTrials_max, dim = c(nPpts, nTimes))
for (i in 1:nPpts) {
  for (t in 1:nTimes) {
  internalChosen_neg[i,t,] <- with(data_long_all, internalChosen[ID==i & sess==t & neg_pos==0])
  internalChosen_pos[i,t,] <- with(data_long_all, internalChosen[ID==i & sess==t & neg_pos==1])
  globalChosen_neg[i,t,] <- with(data_long_all, globalChosen[ID==i & sess==t & neg_pos==0])
  globalChosen_pos[i,t,] <- with(data_long_all, globalChosen[ID==i & sess==t & neg_pos==1])
  }
}

## create list to pass to stan
data_list = list(
  nTimes = nTimes,
  nPpts = nPpts,
  nTrials_max = nTrials_max/2,         # max number of trials per  session per participant
  nT_ppts = nT_ppts,                   # actual number of trials per session per participant
  condition = int_conds$condition01,   # 0 = control, 1 = psychoed
  internal_neg = internalChosen_neg,
  internal_pos = internalChosen_pos,
  global_neg = globalChosen_neg,
  global_pos = globalChosen_pos
)

## fit model using rstan
fit <- stan(
  file = paste0("./stan-models/", model, ".stan"),
  data = data_list,
  chains = 4,               # run 4 separate chains to assess convergence
  warmup = 1000,            # these are used to tune the sampler and ’burn in’
  iter = 2000,              # number of iterations (#kept = chains*(iter - warmup))
  cores = nCores            # chains to be run in parallel on separate cores (if possible)
)

## save
saveRDS(fit, file = paste0("./stan-fits/", model ,"-", task_ver, "-fit.rds"))
# ## OR load presaved models
# fit <- readRDS(file = paste0("./stan-fits/", model ,"-", task_ver, "-fit.rds"))

# summary of sampling diagnostics
check_hmc_diagnostics(fit)

# plot pairs of sampling distributions for an example participant
pairs(fit, pars=c("theta_internal_neg[1,1]", "theta_internal_neg[2,1]",
                  "theta_global_neg[1,1]",   "theta_global_neg[2,1]"))
pairs(fit, pars=c("theta_internal_pos[1,1]", "theta_internal_pos[2,1]",
                  "theta_global_pos[1,1]",   "theta_global_pos[2,1]"))

# plot intervention and individual t2 a[irs ]
pairs(fit, pars=c("theta_int_internal_neg", "theta_int_global_neg"))
pairs(fit, pars=c("theta_int_internal_pos", "theta_int_global_pos"))
```
  
```{r int_IG_plot2}
# test-retest posteriors
R_theta_neg <- as.data.frame(summary(fit, pars = c("R_theta_neg"))$summary)
R_theta_pos <- as.data.frame(summary(fit, pars = c("R_theta_pos"))$summary)

# print mean and mean se for each parameter
# across parameters
print(paste0("R for internal-global attributions for negative items at t1 (mean): ", 
             round(R_theta_neg$mean[2],2)), quote=FALSE)
print(paste0("R for internal-global attribtutions for negative items at t1 (mean SE): ",
             round(R_theta_neg$se_mean[2],3)), quote=FALSE)

print(paste0("R for internal-global attributions for positive items at t1 (mean): ", 
             round(R_theta_pos$mean[2],2)), quote=FALSE)
print(paste0("R for internal-global attribtutions for positive items at t1 (mean SE): ",
             round(R_theta_pos$se_mean[2],3)), quote=FALSE)

print(paste0("R for internal-global attributions for negative items at t2 (mean): ", 
             round(R_theta_neg$mean[12],2)), quote=FALSE)
print(paste0("R for internal-global attribtutions for negative items at t2 (mean SE): ",
             round(R_theta_neg$se_mean[12],3)), quote=FALSE)

print(paste0("R for internal-global attributions for positive items at t2 (mean): ", 
             round(R_theta_pos$mean[12],2)), quote=FALSE)
print(paste0("R for internal-global attribtutions for positive items at t2 (mean SE): ",
             round(R_theta_pos$se_mean[12],3)), quote=FALSE)

# across time points
print(paste0("R for internal attributions for negative items t1-t2 (mean): ", 
             round(R_theta_neg$mean[3],2)), quote=FALSE)
print(paste0("R for internal attributions for negative items t1-t2 (mean SE): ",
             round(R_theta_neg$se_mean[3],3)), quote=FALSE)

print(paste0("R for internal attributions for positive items t1-t2 (mean): ", 
             round(R_theta_pos$mean[3],2)), quote=FALSE)
print(paste0("R for internal attributions for positive items t1-t2 (mean SE): ",
             round(R_theta_pos$se_mean[3],3)), quote=FALSE)

print(paste0("R for global attributions for negative items t1-t2 (mean): ", 
             round(R_theta_neg$mean[8],2)), quote=FALSE)
print(paste0("R for global attributions for negative items t1-t2 (mean SE): ",
             round(R_theta_neg$se_mean[8],3)), quote=FALSE)

print(paste0("R for global attributions for positive items t1-t2 (mean): ", 
             round(R_theta_pos$mean[8],2)), quote=FALSE)
print(paste0("R for global attributions for positive items t1-t2 (mean SE): ",
             round(R_theta_pos$se_mean[8],3)), quote=FALSE)

# plot mean param ests session 1 vs 2
posts <- as.data.frame(summary(fit,
                       pars = c("p_internal_pos", "p_internal_neg",
                                "p_global_pos", "p_global_neg"))$summary) %>%
  dplyr::select(mean, sd) %>%
  add_rownames(var = "var") %>%
  separate(var, sep="\\[|\\,|\\]", into=c("parameter", "ID", "session"),
           remove=TRUE, extra="drop") %>%
  separate(parameter, sep=-3, into=c("parameter", "item_valence")) %>%
  mutate(parameter = sub("l_", "l", parameter),
         parameter = factor(parameter, levels = c("p_internal", "p_global")))

## plot
p1 <- posts %>%
  pivot_wider(id_cols=c("ID", "parameter", "item_valence"),
              names_from = "session", values_from = c("mean", "sd")) %>%
  mutate(condition = ifelse(ID %in% control_IDs, "control", "restructuring")) %>%
  ggplot(aes(x=mean_1, y=mean_2, group=condition, colour=condition)) +
  geom_abline(slope = 1, linetype="dashed", colour="grey") +
  geom_point() +
  geom_errorbarh(aes(xmin = mean_1-sd_1, xmax = mean_1+sd_1), alpha=.4) +
  geom_errorbar(aes(ymin = mean_2-sd_2, ymax = mean_2+sd_2), alpha=.4) +
  geom_smooth(method = "lm", se=FALSE, formula = y ~ x) +
  scale_colour_manual(values=colours3) + 
  scale_fill_manual(values=colours3) +
  labs(x = "mean (sd) time 1", y ="mean (sd) time 2") +
  theme_minimal() + facet_grid(cols=vars(item_valence), rows=vars(parameter)) + labs()  + 
  theme(aspect.ratio=4/3.5)
print(p1)
# ggsave(filename = paste0("./figures/", task_ver, "-params-by-time.svg"),
#        plot = last_plot(), device = "svg", dpi = 300)

# plot posteriors for effects of interest using tidybayes  (quick and pretty!)
fit_tidy <- fit %>% 
  gather_draws(`mu_internal_theta_neg[1]`, `mu_internal_theta_neg[2]`, 
               `mu_internal_theta_pos[1]`, `mu_internal_theta_pos[2]`,
               `mu_global_theta_neg[1]`,   `mu_global_theta_neg[2]`, 
               `mu_global_theta_pos[1]`,   `mu_global_theta_pos[2]`, 
               theta_int_internal_neg, theta_int_internal_pos,
               theta_int_global_neg,   theta_int_global_pos) %>%
  mutate(var_type = ifelse(grepl("_int_", .variable), "intervention effect", "group mean"),
         var_type = factor(var_type, levels = c("group mean", "intervention effect")),
         .variable = factor(.variable, levels = c(
           "mu_internal_theta_neg[1]", "mu_internal_theta_neg[2]",
           "mu_internal_theta_pos[1]", "mu_internal_theta_pos[2]",
           "mu_global_theta_neg[1]",   "mu_global_theta_neg[2]",
           "mu_global_theta_pos[1]",   "mu_global_theta_pos[2]",
           "theta_int_internal_neg", "theta_int_internal_pos",
           "theta_int_global_neg", "theta_int_global_pos")))

p <- fit_tidy %>%
  ggplot(aes(y = fct_rev(.variable), x = .value, fill = var_type)) +
  stat_gradientinterval(.width = c(.9, .5),  slab_size = 1) +
  scale_fill_manual(values = colours2) +
  geom_vline(xintercept = 0, colour = "grey") + theme_minimal() +
  theme(legend.position = "none") + theme(aspect.ratio=4/3) + labs(x="", y="")
p
# ggsave(filename = paste0("./figures/", task_ver, "-causal-attr-task-means-ints-CIs-gradient.svg"),
#        plot = last_plot(), device = "svg", dpi = 300)

# print numerical values for posteriors (raw / untransformed values)
params90cis <- summary(fit, pars = c("mu_internal_theta_neg[1]", "mu_internal_theta_neg[2]",
                                     "mu_internal_theta_pos[1]", "mu_internal_theta_pos[2]",
                                     "mu_global_theta_neg[1]",  "mu_global_theta_neg[2]",
                                     "mu_global_theta_pos[1]",  "mu_global_theta_pos[2]",
                                     "theta_int_internal_neg", "theta_int_internal_pos",
                                     "theta_int_global_neg", "theta_int_global_pos"), 
                       probs = c(0.05, 0.95))$summary
print(params90cis)

## new plotting (standardized / transformed)
## re-transform group means to output probabilities (cf raw theta estimates which control this)
## and convert intervention effects to ~SMDs
# theta_x_int_std =theta_x_int_sd / sqrt(sigma_theta[2])
# first, get posterior (pooled) variance estimates for theta_x at time 2 (internal, global)
params90cis <- summary(fit, pars = c("pars_sigma_neg[3]",
                                     "pars_sigma_neg[4]", 
                                     "pars_sigma_pos[3]",
                                     "pars_sigma_pos[4]"), probs = c(0.05, 0.95))$summary
sigma_theta_int_neg_t2 <- params90cis[1,1]
sigma_theta_glob_neg_t2 <- params90cis[2,1]
sigma_theta_int_pos_t2 <- params90cis[3,1]
sigma_theta_glob_pos_t2 <- params90cis[4,1]

fit_tidy2 <- fit_tidy %>%
  mutate(.value2 = case_when(.variable =="mu_internal_theta_neg[1]" ~ inv.logit(.value),
                             .variable =="mu_internal_theta_neg[2]" ~ inv.logit(.value),
                             .variable =="mu_internal_theta_pos[1]" ~ inv.logit(.value),
                             .variable =="mu_internal_theta_pos[2]" ~ inv.logit(.value),
                             .variable =="mu_global_theta_neg[1]" ~ inv.logit(.value),
                             .variable =="mu_global_theta_neg[2]" ~ inv.logit(.value),
                             .variable =="mu_global_theta_pos[1]" ~ inv.logit(.value),
                             .variable =="mu_global_theta_pos[2]" ~ inv.logit(.value),
                             .variable =="theta_int_internal_neg" ~ .value/sqrt(sigma_theta_int_neg_t2),
                             .variable =="theta_int_internal_pos" ~ .value/sqrt(sigma_theta_int_pos_t2),
                             .variable =="theta_int_global_neg"   ~ .value/sqrt(sigma_theta_glob_neg_t2),
                             .variable =="theta_int_global_pos"   ~ .value/sqrt(sigma_theta_glob_pos_t2),
                             TRUE ~ .value))
p2 <- fit_tidy2 %>%
  ggplot(aes(y = fct_rev(.variable), x = .value2, fill = var_type)) +
  stat_gradientinterval(.width = c(.9, .5),  slab_size = 1) +
  scale_fill_manual(values = colours2) +
  geom_vline(xintercept = 0, colour = "grey") + theme_minimal() +
  theme(legend.position = "none") + theme(aspect.ratio=4/3) + labs(x="", y="")
p2
# ggsave(filename = paste0("./figures/", task_ver, "-means-ints-CIs-gradient-transf-smd.svg"),
#        plot = last_plot(), device = "svg", dpi = 300)
```



